---
title: "Extracting Fama French Data"
description: |
  Using R to extract all the dataset from Kennith French's Darthmouth website.
author:
  - name: Taylor Wood
    url: https://twood.blog/famafrenchextract
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
The following code is used to download all of the data on Kenneth R. French's Website.  

First we need to include the necessary packages.  

```{r dependencies, echo=TRUE}
library(tidyverse)
library(rvest)
library(readr)
library(lubridate)
library(janitor)
```



Next we need to download all the links to the CSV files in 
on Kenneth R. French's website.  

```{r downloadLinks, echo=TRUE}
# Kenneth French's Data Library Webpage
url <- "https://mba.tuck.dartmouth.edu/pages/faculty/ken.french/data_library.html"

# Read HTML and find <b> nodes
# Look for nodes with "CSV.zip" 
csvLinks <- url %>%
  read_html() %>%
  html_nodes("b") %>%
  as.character() %>%
  as_tibble(column_name = "value") %>%
  mutate(`lead2` = lead(`value`, 2)) %>%
  filter(str_detect(`lead2`, "CSV.zip")) %>%
  transmute(
    `Name` = str_extract(`value`, "(?<=\\<b\\>).*(?=\\</b\\>)"),
    `Link` = str_extract(`lead2`, "ftp.+zip"))
```



Once we have Data Frame of the all the CSV file names we need to download the
zipped files, unzip them, and store them in a list.  

```{r donwloadRawData, echo=TRUE}
FFExtract <- function(df){
  
  # Find Cells with Name/Description and Extract
  nameLoc <- str_which(df$V1, "[A-z]+")
  if(length(nameLoc) == 0){
    name <- NA
  }else{
    name <- str_c(str_trim(df[nameLoc,1][[1]]), collapse = " ")
  }
  
  # Find Rows with Data and Extract
  dataLoc <- str_which(df$V1, "^\\s*[0-9]{4,}")
  data <- df[dataLoc,] %>%
    `colnames<-`(df[min(dataLoc)-1,]) %>%
    clean_names() %>%
    rename(`Date` = 1) %>%
    mutate_all(as.numeric)
  
  return(list("Description" = name, "Data" = data))
}

# Function to Download CSV Data from Child
dataDownload <- function(name, child){
  
  # Create working URL of child 
  fileName <- child %>%
    str_extract("(?<=/).+(?=\\_CSV\\.zip)") %>%
    str_c(".CSV")
  
  # Create tempfile for download
  temp <- tempfile()
  
  # Download zip file from Ken French website
  download.file(
    url = str_c(
      "https://mba.tuck.dartmouth.edu/pages/faculty/ken.french/",
      child),
    destfile = temp)
  
  # Unzip data file
  unzip(temp)
  
  # Read in CSV file
  rawData <- read.csv(
    fileName,
    header = FALSE,
    sep = ",",
    col.names = paste0("V",seq_len(200)),
    fill = T,
    blank.lines.skip = F)
  
    # Remove Leading rows
  Desc <- rawData[1:(min(str_which(rawData$V1, "^[0-9]{4,}"))-2),1]
  rawData <-rawData[(min(str_which(rawData$V1, "^[0-9]{4,}"))-2):nrow(rawData),]
  
  # Split data into different data chunks
  Data <- rawData %>% 
    filter(!str_detect(.[[1]], "Copyright")) %>%
    select_if(~!all(is.na(.))) %>%
    select_if(~!all(. == "")) %>%
    mutate(
      `istext` = ifelse(str_detect(.[[1]], "[a-z]+"),1,0),
      `change` = ifelse(`istext` == 1 & lag(`istext`,1, default = 0) == 0 , 1, 0),
      `group` = cumsum(`change`)) %>%
    select(-c(`istext`, `change`)) %>%
    group_split(`group`, .keep = F)

  extractedList <- lapply(Data, FFExtract)

  return(list("Description" = Desc,
              "Data" = extractedList))
}
```

Now we can download all of the data.  

```{r, echo=TRUE}
FamaFrenchData <- mapply(dataDownload, 
                         name = csvLinks$Name,
                         child = csvLinks$Link,
                         SIMPLIFY = FALSE)
```

From there we can use the data for further analysis.  

```{r, echo = TRUE}
FF3 <- FamaFrenchData$`Fama/French 3 Factors [Daily]`$Data[[1]]$Data %>%
  transmute(
    `Date` = as.Date(as.character(`Date`), "%Y%m%d"),
    `Rm_Rf_Cum` = cumprod(1 + `mkt_rf`/100) - 1,
    `SMB_Cum` = cumprod(1 + `smb`/100) - 1,
    `HML_Cum` = cumprod(1 + `hml`/100) - 1) %>%
  pivot_longer(
    cols = c(`Rm_Rf_Cum`, `SMB_Cum`, `HML_Cum`),
    names_to = "factor_name",
    values_to = "factor_value")

ggplot(FF3) +
  aes(x = Date, y = factor_value, colour = factor_name) +
  scale_y_continuous(labels = scales::percent) +
  geom_line(size = 0.5) +
  scale_color_viridis_d(option = "viridis", direction = 1) +
  labs(
    x = "Date",
    y = "Cumualtive Return",
    title = "Fama French 3 Factor Model",
    subtitle = "Cumulative Return",
    color = "Factor Name"
  ) +
  theme_minimal()
```
